Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-07-14 doc/paper.tex
oops, forgot the transduction grammar and prereordering comments. running out of time...

% TODO: corpus generation is (almost like?) a sub-problem
% TODO: a reference to ``blessing of dimensionality'' for the inverse problem? how it's relatively easy to generate an exponentially large number of sentences?
% TODO: tree-based methods become a lot easier too? I even construct a random stupid tree myself...
Other practical considerations shine a favorable light on parallel corpus generation.
As the ``inverse'' of translation, multilingual generation is also an easier task in some senses.
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would ``know'' {\em a priori} whether it is writing a sentence about, say, (financial) banks or (river) banks.
} segmentation of languages like Chinese \mycitep{segmenter:2008} becomes a non-issue, and there is the potential for generating annotated corpora with zero annotation error.
Parallel corpora also provide a natural avenue to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems. %, as discussed in Section \ref{sec:related}. % that's not really discussed there anymore, is it?
And unlike the various different MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\noindent\makebox[\linewidth]{\rule{\paperwidth}{0.4pt}} % full-width horizontal line
%TODO: detailed system description
% code section - do this first, since it MAY shed light on what linguistic phenomena are easy to implement?
    % oh yes, definitely, because then you can explain things briefly in terms of their implementation
%
% Simplistic Linguistics
% maybe use a bulleted list for the different linguistic phenomena supported?
%Clause: transitive, meta, modal
%basic support for NP, ADJP, ADVP (targeting ADJP or Clause), pronouns with antecedents 
%prepositions, which are really relations, allowing semantic constraints on both object and target
%indirect object as a special case of ``null preposition''
%- not necessarily linguistically well-founded, but good enough
%past tenses via transformation of present tense sentence
% other transformations: topicalization, and some hokier stuff like \zh{把} 
%Custom (``template-based mode''): lots of iterals, maybe mark an NP somewhere

@@

% preemptively answering reviewer question: yeah yeah, the trained tmcombine weights are like 0.001 for my generated corpora
The results of these experiments are shown in Table~\ref{tab:results}.
For the table on the left, note that adding in-domain data improves performance significantly more than a much larger amount of out-of-domain data (MultiUN), a commonly reported phenomenon \mycitep{tmcombine}.
TED talks, which are prepared lectures, fall somewhere between the conversational language of movie subtitles and the formal written language of UN resolutions.
As such, the OpenSubtitles\footnote{http://www.opensubtitles.org/} 2016 corpus \mycitep{opensubtitles:2016} from the OPUS project was also out-of-domain; results for this corpus were similar to MultiUN and are not shown.
For the table on the right, note the diminishing returns of out-of-domain data as the amount of in-domain data increases.
These observations underscore the importance of domain overlap and hint at the potential value to be found in the amplification of in-domain corpora.

@@

In tables with MultEval scores below, metric scores for all systems were generated using MultEval 0.5.1 \mycitep{multeval}: jBLEU V0.1.1 (an exact reimplementation of NIST's mteval-v13.pl without tokenization); Translation Error Rate (TER) V0.8.0. 
The $p$-values are relative to the baseline and indicate whether a difference of this magnitude (between the baseline and the system on that line) is likely to be generated again by some random process (a randomized optimizer). 
Metric scores are averages over three runs per system. 
$s_{sel}$ indicates the variance due to test set selection and has nothing to do with optimizer instability.\footnote{
    This paragraph and the tables with MultEval results are adopted from \LaTeX~code automatically generated by MultEval.
}

@@

%\pagebreak %a man a plan a canal a panama. yes, it looks like ANY new reference will cause AMTA to go over limit.

>>>>

% TODO: corpus generation is (almost like?) a sub-problem
% TODO: a reference to ``blessing of dimensionality'' for the inverse problem? how it's relatively easy to generate an exponentially large number of sentences?
% TODO: tree-based methods become a lot easier too? I even construct a random stupid tree myself...
Other practical considerations shine a favorable light on parallel corpus generation.
As the ``inverse'' of translation, multilingual generation is also an easier task in some senses.
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would ``know'' {\em a priori} whether it is writing a sentence about, say, (financial) banks or (river) banks.
} segmentation of languages like Chinese \mycitep{segmenter:2008} becomes a non-issue, and there is the potential for generating annotated corpora with zero annotation error.
In principle, high-quality alignments could be obtained at the word or phrase level for use in the statistical MT process, and these alignments 
could also be potentially useful in training methods for MT subtasks such as syntactic pre-reordering \mycitep{prereordering}.
Parallel corpora also provide a natural avenue to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems. %, as discussed in Section \ref{sec:related}. % that's not really discussed there anymore, is it?
And unlike the various different MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\noindent\makebox[\linewidth]{\rule{\paperwidth}{0.4pt}} % full-width horizontal line
%TODO: detailed system description
% code section - do this first, since it MAY shed light on what linguistic phenomena are easy to implement?
    % oh yes, definitely, because then you can explain things briefly in terms of their implementation

%Language-specific rules (e.g., word order) are confined to ad hoc bits of code the generator module.\footnote{
%    Formally, this might be considered an extremely rudimentary synchronous (bilingual) context-free grammar of sorts \mycitep{transduction}, one that also admits long-range dependencies and semantic constraints.
%    But it would be a disservice to those other well-developed methods to refer to our system as anything other than an ad hoc way to generate multilingual sentences.
%}

%
% Simplistic Linguistics
% maybe use a bulleted list for the different linguistic phenomena supported?
%Clause: transitive, meta, modal
%basic support for NP, ADJP, ADVP (targeting ADJP or Clause), pronouns with antecedents 
%prepositions, which are really relations, allowing semantic constraints on both object and target
%indirect object as a special case of ``null preposition''
%- not necessarily linguistically well-founded, but good enough
%past tenses via transformation of present tense sentence
% other transformations: topicalization, and some hokier stuff like \zh{把} 
%Custom (``template-based mode''): lots of iterals, maybe mark an NP somewhere

@@

% preemptively answering reviewer question: yeah yeah, the trained tmcombine weights are like 0.001 for my generated corpora
The results of these experiments are shown in Table~\ref{tab:results}.
For the table on the left, note that adding in-domain data improves performance significantly more than a much larger amount of out-of-domain data (MultiUN), a commonly reported phenomenon \mycitep{tmcombine}.
TED talks, which are prepared lectures, fall somewhere between the conversational language of movie subtitles and the formal written language of UN resolutions.
As such, the OpenSubtitles\footnote{http://www.opensubtitles.org/} 2016 corpus \emnlponly{\mycitep{opensubtitles:2016}} from the OPUS project was also out-of-domain; results for this corpus were similar to MultiUN and are not shown. % AMTA: removed citation to keep reference list under length
For the table on the right, note the diminishing returns of out-of-domain data as the amount of in-domain data increases.
These observations underscore the importance of domain overlap and hint at the potential value to be found in the amplification of in-domain corpora.

@@

In tables with MultEval scores below, metric scores for all systems were generated using MultEval 0.5.1 \mycitep{multeval}: jBLEU V0.1.1 (an exact reimplementation of NIST's mteval-v13.pl without tokenization); Translation Error Rate (TER) V0.8.0. 
The $p$-values are relative to our baseline and indicate whether a difference of this magnitude (between the baseline and the system on that line) is likely to be generated again by some random process (a randomized optimizer). 
Metric scores are averages over three runs per system. 
$s_{sel}$ indicates the variance due to test set selection and has nothing to do with optimizer instability.\footnote{
    This paragraph and the tables with MultEval results are adopted from \LaTeX~code automatically generated by MultEval.
}

@@

\pagebreak
